---
title: "ECON ANALYSIS"
output: html_notebook
---

Base Theory and Loation Quotient

GOALS:
  1 - Define Base Theory
  2 - Apply Base Theory
  3 - Calculate LQ and Basic Employment
  4 - Shift Share Economic Method
  
  Want to be able to define how various economies and industries are changing
  and how.
  
  Location Quotient (LQ): What does an area specialize in
  
  Econ Shift Share: What are the shifts in specialization over time
  
  Base Theory: The economy is made up as two components
    1: Basic Sector - Export or Non-Local Sector. Products produced in an area
        that are then sold elsewhere. Therefore income comes into the economy
    2: Non-Basic - Sells within the region and recirculates income. Exists to serve
        the basic economy

  Economic Base Multiplier (BM)
    How much employment is generated from the creation of 1 basic sector job
    
    We assume BM is constant and can be projected forward.
    
    BM = e_t/b_t

    Govt. is considered basic employment as it brings in money that supports
    non basic services in an are.
    
    To determine econmic base we must calculate the number of jobs in an area. Sometimes
    declining jobs is not always a sign of a declining sector.
    
    LQ Analysis
      If an economy specializes in a specific sector over the basline of the reference economy
        then a reference will show as a value greater than 1. If LQ = 1 all demand is being
        met in the local economy. If LQ > 1 then outputs are being exported and jobs related
        to those export is calculated for the amount over 1. 
      Compare local production in an industry to number of jobs in a reference economy
      LQ = local share / national share
      
      LQ_i = (e_i/e) / (E_i/E)
      
      e_lq_1 = e*(E_i/E) this gets us to LQ = 1
      
      basic employment = e_i - e_lq_i = e_i = (e*(E_i/E))
      
      Assumes that local economy reflects the performance and behavior of reference economy
      Assumes that there is constant demand across space. There are variations in local demand
        that do not match reference economy (.: LQ > 1 may not indicate an export)

        
```{r}
naics_data <- read.csv('sf_location_quotient_raw.csv')
naics_data 
```

e = Employment in a sector in a region
E = Employmnet in a sector national

i = industry

```{r}

naics_data$lq_i <- (naics_data$sf_share/naics_data$sf_share[1])/(naics_data$us_share/naics_data$us_share[1])

i <- 1
print(length(naics_data$industry))
for (i in 1:length(naics_data$industry)) {
    print(naics_data$lq_i[i])
    if (is.na(naics_data$lq_i[i])) {
      naics_data$basic[i] <- 0
      naics_data$n_basic[i] <- 0
    }
    else if (naics_data$lq_i[i] > 1) {
      #print("I am in 1")
      naics_data$basic[i] <- naics_data$sf_val[i]-(naics_data$sf_val[1]*(naics_data$us_val[i]/naics_data$us_val[1]))
      naics_data$n_basic[i] <- naics_data$sf_val[i] - naics_data$basic[i]
    }
    else {
      #print("I am in 2")
      naics_data$basic[i] <- 0
      naics_data$n_basic[i] <- naics_data$sf_val[i]
    }
    i = i + 1
}

naics_data

b_sum = sum(naics_data$basic[2:length(naics_data$basic)])
n_sum = sum(naics_data$n_basic[2:length(naics_data$n_basic)])
basic_mul = n_sum/b_sum

basic_mul
```

