---
title: "Projection Function Documentation"
output: html_notebook
---

First you must load the environment necessary to perform the analysis. This includes loading the 2 .txt files
that contain the necessary data.

This notebook allows for the grader or student to access and manipulate the functions as a means of analyzing
or grading them.

```{r}
library(nls2)

ga_counties <- read.csv('a01-gareis2003.txt')
ga_proj <- read.csv('a40-gaprojected.txt')

analysis_level <- data.frame("FIPS" = unique(ga_counties$FIPS), "AREANAME" = unique(ga_counties$AREANAME))
analysis_type <-  data.frame("LINECODE" = unique(ga_counties$LINECODE), "LINETITLE" = unique(ga_counties$LINETITLE))
analysis_years <- data.frame(unique(ga_counties$YEAR))

default_base <- 1969
default_launch <- 2003
default_proj <- 2010

#DEFAULT STATE:
# Level - FIPS: 13001, Appling County GA
# Type - LINECODE: 20, Population
default_level <- 13001
default_type <- 20

a <- data.frame( pop = subset(ga_counties, FIPS == default_level & LINECODE == default_type)$VALUE,
                 year = subset(ga_counties, FIPS == default_level & LINECODE == default_type)$YEAR-1968)
a <- a[with(a, order(year)),]
a_j <- data.frame( pop = subset(ga_counties, FIPS == 13000 & LINECODE == default_type)$VALUE,
                 year = subset(ga_counties, FIPS == 13000 & LINECODE == default_type)$YEAR-1968)
a_j <- a_j[with(a, order(year)),]
b <- data.frame(subset(ga_proj, Year == default_proj & DataType == 'Population'))

x = data.frame(a$year)
y = data.frame(a$pop)

```

Models must be declared and constructed

LINEAR MODELS:
To construct these models the lm (linear model builder) is used to construct a simple polynomial
linear fit. a+bx, a+b_1x+b_2x^2, ..., a+b_1x+b_2x^2+...+b_n-1x^n-1+b_nx^n the model is buildt
and a prediction run from a time t+1 out to the projection distance (p_d).

```{r}
  #Linear
  linear_proj <- function (x, y, t_l, p_d) {
    lin.mod <- lm(y ~ x)
    lin.pre <- predict(lin.mod, newdata = data.frame(x = seq(t_l+1,t_l+p_d)))
    return(lin.pre)
  }
  
  #Quadratic
  quad_proj <- function (x, y, t_l, p_d) {
    qua.mod <- lm(formula = y ~ poly(x,2))
    qua.pre <- predict(qua.mod, newdata = data.frame(x = seq(t_l,t_l+p_d)))    
    return(qua.pre)
  }

  #Cubic
  cube_proj <- function(x, y, t_l, p_d) {
    cub.mod <- lm(formula = y ~ poly(x,3))
    cub.pre <- predict(cub.mod, newdata = data.frame(x = seq(t_l,t_l+p_d)))
    return(cub.pre)
  }
```

NONLINEAR MODELS
These models are constructed from the nls (non linear least squares model builder). Given that
the data has very small residuals nls2 was used to brute force the finding of the method. This allows
for the construction of geometric, expoential and logistic growth models.

The method for assembly is as follows

(1) Lineraize the model to establish residual values  from which to estimate initial fitting values
  Lineraized models, as the name suggests simplify the model into some form of log(y) ~ x
    Simple Exponential: log(y)/log(y_b) ~ x
    Simple Geometric:   log(y) - log(y_b) ~ x
    Modified Expoential:log(y/y_b) ~ x
    Logistic:           log(1/y-c) ~ x || c = 1/(1.5*y_l)
(2) Create nls2 model and projections

Projections are run from 1 to a number of years, p_d, from launch. i.e. 2010 = 7.
p_d is calculated as projection_year - launch_year.

The carrying capacity is considered to be 1.5 * y_l as instructions indicate.

```{r}
  #Simple Exponential
  s_exp <- function(x, y, t_l, p_d) {
    s_e.0 <- lm(log(y[2:length(y)]/log(y[1])) ~ x[2:length(x)])
    start <- list(a=exp(coef(s_e.0)[1]), b=coef(s_e.0)[2], c = y[length(y)]*1.5)
  
    s_e.mod <- nls2(y ~ a * exp( b * x ), data = data.frame(y), start = start, algorithm = "brute-force")
    s_e.pre <- predict(s_e.mod, newdata = data.frame(x = seq(1,p_d)))    
    return(s_e.pre)
  }

  #Simple Geometric
  s_g <- function(x, y, t_l, p_d) {
    s_ge.0 <- lm((log(y[2:length(y)]-(y[1])) ~ x[2:length(x)]))
    start <- list(a=exp(coef(s_ge.0)[1]), b=coef(s_ge.0)[2], c = y[length(y)]*1.5)
    
    s_ge.mod <- nls2(y ~ a * (1 + b)^x + c, data = data.frame(y), start = start, algorithm = "brute-force")
    s_ge.pre <- predict(s_ge.mod, newdata = data.frame(x = seq(1,p_d))) 
    return(s_ge.pre)
  }
  
  #Modified Exponential
  m_e <- function(x, y, t_l, p_d) {
    m_e.0 <- lm(log(y[2:length(y)]/log(y[1])) ~ x[2:length(x)])
    start <- list(a=exp(coef(m_e.0)[1]), b=coef(m_e.0)[2], c = y[length(y)]*1.5)
    
    m_e.mod <- nls2(y ~ a * exp( b * x ), data = data.frame(y), start = start, algorithm = "brute-force")
    m_e.pre <- predict(m_e.mod, newdata = data.frame(x = seq(1,p_d)))   
    return(m_e.pre)
  }
  
  #Logistic
  lo <- function(x, y, t_l, p_d) {
    c.0 = 1/(1.5*y[length(y)])
    lo.0 <- lm(log((1/y) - c.0) ~ x)
  
    summary(lo.0)
    start <- data.frame(a = exp(coef(lo.0)[1]), b = exp(coef(lo.0)[2]), c = c.0)
  
    lo.mod <- nls2(y ~ 1/(c+a*(b^x)), data.frame(y), start = start, algorithm = "brute-force")
    lo.pre <- predict(lo.mod, newdata = data.frame(x = seq(1,p_d))) 
    return(lo.pre)
  }
```

RATIO PROJECTION METHODS
Due to challenges in creating functioning models for the ratio projection methods these
methods have been solved numerically per the equations given in class.
  Constant Share: P_it = (P_il/P_jl)*P_jt
  Shift Share: P_it = [(P_il/P_jl)+((P_il/P_jl)-(P_ib/P_jb))*(1/(t_l-t_b))*z]
  Growth Share: P_it = P_il+[(P_il-P_ib)/(P_jl- P_jb)]*(P_jt-P_jl)

Since these equations are solved numerically these functions are not graphed nor is any attempt to graph them 
made as any part of this project due to challenges with R and graphing this kind of non-modeled data in a complex
visualization.

```{r}
  #Constant Share
  c_s <- function(a, a_j, b) {
    c_sp <- (a$pop[length(a$pop)]/a_j$pop[length(a_j$pop)])*b$Value
    return(c_sp)
  }
  #Shift Share
  s_s <- function (a, a_j, b, p_d){
    s_l <- a$pop[length(a$pop)]/a_j$pop[length(a_j$pop)] #launch share
    d_s <- (a$pop[length(a$pop)]/a_j$pop[length(a_j$pop)])-(a$pop[1]/a_j$pop[1]) #difference of share
    i_t <- 1/a$year[length(a$year)] #1/time
    
    s_sp <- (s_l+(d_s*i_t*p_d))*b$Value
    return(s_sp)
  }
  #Share of Growth
  s_gr <- function (a, a_j, b){
    d_i <- a$pop[length(a$pop)] - a$pop[1]
    d_j <- a_j$pop[length(a_j$pop)] - a_j$pop[1]
    d_t <- b$Value - a_j$pop[length(a_j$pop)]
    
    s_gp <- a$pop[length(a$pop)] + (d_i/d_j)*(d_t)
    return(s_gp)
  }
```

```{r}
te_l <- linear_proj(a$year, a$pop, a$year[length(a$year)], 10)
te_q <- quad_proj(a$year, a$pop, a$year[length(a$year)], 10)
te_c <- cube_proj(a$year, a$pop, a$year[length(a$year)], 10)

te_se <- s_exp(a$year, a$pop, a$year[length(a$year)], 10)
te_sg <- s_g(a$year, a$pop, a$year[length(a$year)], 10)
te_me <- m_e(a$year, a$pop, a$year[length(a$year)], 10)
te_lo <- lo(a$year, a$pop, a$year[length(a$year)], 10)

te_cs <- c_s(a, a_j, b)
te_ss <- s_s(a, a_j, b, 10)
te_sgr <- s_gr(a, a_j, b)

te_l[length(te_l)]
```
 
```{r}
ggplot(a, aes(a$year, a$pop)) + geom_point(aes(size = a$pop))
```
